---
title: "microPEM First Look"
output: pdf_document
---

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide', fig.align='center', fig.height= 10, fig.width=7.5}


#MicroPEM upload and inital data check
#microPEM files collected in Ghana

#Batch upload of microPEM files

library(plyr)
library(lubridate)
library(ggplot2)
library(dplyr)
setwd("/Applications/Rdata/Ghana_microPEM/Colo_test") # change for correct file path

#Individual microPEM tests
#Upload all Files
indTest <- list.files(".", recursive = T, pattern = "^(UGF320414N)", full.names = T) #can edit pattern based on file name you want. '
length(indTest)

#create columns with Sample ID, Unit ID, and Filter ID from file name
#combine date and time columns
pem.import <- function(PEM){
  indtest_01 <- read.csv(PEM, skip = 28, stringsAsFactors = F, header = T, row.names = NULL)  
  names(indtest_01) <- c("Date", "Time","PM", "Temp",	"RH",	"Battery",	"InletPress",	"FlowOrificePress",	"Flow",	"Xaxis",	"Yaxis",	"Zaxis",	"VectorSumComposite", "action", "Wearing Compliance", "ValidityWearingCompliance validation")
  filename <- PEM
  filename2 <- sub(".csv", "", filename)
  filename2 <- sub("./", "", filename2)
  filename_vec <- strsplit(filename2, split = "_")[[1]]
  indtest_01$UnitID <- filename_vec[1] 
  indtest_01$FilterID <- filename_vec[2]
  indtest_01$SampleID <- filename_vec[3]
  indtest_01$Date <-dmy(indtest_01$Date)
  indtest_01$Datetime <- as.POSIXct(paste(indtest_01$Date, indtest_01$Time), origin="1970-01-01",format = "%Y-%m-%d %H:%M:%S",tz='Africa/Accra')
  indtest_01}


names(indTest) <- indTest

indTest_stack <- ldply(indTest, pem.import, .progress = "text")

#check for low bat stop 
batt <- filter(indTest_stack, action == "Low Battery Stop")

#Create minute averages data table
indTest_stack$Time_minute <- floor_date(indTest_stack$Datetime, unit = "minute")

indtest_minavg <- indTest_stack %>% group_by(UnitID, FilterID, SampleID, Time_minute) %>% summarise(Date[1], mean(PM, na.rm = TRUE), mean(Temp, na.rm = TRUE),
                                                                                  mean(RH, na.rm = TRUE),mean(Battery, na.rm = TRUE), mean(InletPress, na.rm = TRUE),
                                                                                  mean(FlowOrificePress, na.rm = TRUE), mean(Flow, na.rm = TRUE),
                                                                                  mean(VectorSumComposite, na.rm = TRUE), sd(VectorSumComposite, na.rm = TRUE) )

#Change Colmn names
names(indtest_minavg) <- c("UnitID", "FilterID","SampleID", "Time","Date","PM",	"Temp","RH","Battery","InletPress",	"FlowOrificePress",	"Flow",	"VectorSumComposite", "SD_VectorSumComposite")

#### Calculate compliance
#### sd of comp > 0.02 ? Rolling over how many minutes? 
#### used 0.02 as threshold and rolling average of 10 minutes. Change/adjust if needed

indtest_minavg$compliance_thres <- ifelse(indtest_minavg$SD_VectorSumComposite > 0.02, 1, 0)

#rolling mean 

# calculate rollmean compliance
library(zoo)
indtest_minavg$sd_composite_rollmean <- as.numeric(rollapply(indtest_minavg$compliance_thres, width=10,  FUN = mean, align = "center", na.rm = TRUE, fill = NA)) 

indtest_minavg$compliance_rollmean <- ifelse(indtest_minavg$sd_composite_rollmean > 0, 1, 0)

if (sum(!is.na(indtest_minavg$compliance_rollmean)) > 0) {
  indtest_minavg.complete <-  indtest_minavg[complete.cases(indtest_minavg),] 
} else {
  indtest_minavg.complete <- indtest_minavg
}
    


#Creating 24 hour periods. Need to do after minute average to cut timepoints without PM data (i.e when the USB is connected)
no.days <- ceiling(as.numeric(as.duration(indtest_minavg.complete$Time[nrow(indtest_minavg.complete)] - indtest_minavg.complete$Time[1]))/86400) # calculates the difference in time between last and first datetime observation (in seconds), transforms it into days and returns the ceiling of days
    
dayindex <- indtest_minavg.complete$Time[1] + hours(seq(from = 24, to = no.days*24, by = 24))
    
    indtest_minavg.complete$unique_24h <- 1
    
    for (i in 1:no.days) { 
      indtest_minavg.complete$unique_24h <- ifelse ((indtest_minavg.complete$Time > dayindex[i]),  i+1, indtest_minavg.complete$unique_24h)
    } 
    
    

  
####Add HEPA corrections.....

library(changepoint)

Data5 <- subset(indTest_stack, !is.na(PM))
Data4 <- subset(indTest_stack, !is.na(PM))

Data6 = Data5[7:100,]   # exclude the first 6 readings (about 1 minute)
HEPASt = cpt.meanvar(Data6$PM,method="BinSeg", Q=3, minseglen=8)
Data4$HEPAstnumber = HEPASt@cpts[1]        # identify the place where changepoint is
if(Data4$HEPAstnumber==94){                # if there is no changepoint in the first 100 readings then no start HEPA
  Data4$HEPAsttime1 = NA
  Data4$HEPAsttime2 = NA
  Data4$HEPAstvalue1 = NA
  Data4$HEPAstvalue2 = NA
} else {
  Data4$HEPAsttime1 = Data5$Datetime[7]              # the starttime of start HEPA
  Data4$HEPAsttime2 = Data5$Datetime[Data4$HEPAstnumber+6]   # the endtime of start HEPA
  # the mean nephelometer reading in start HEPA
  Data4$HEPAstvalue1 = mean(Data6$PM[1:Data4$HEPAstnumber]) 
  # the mean nephelometer reading after start HEPA period (within the first 100 readings)
  Data4$HEPAstvalue2 = mean(Data6$PM[(Data4$HEPAstnumber+1):94])
}


# ending HEPA
Data7 = Data5[(nrow(Data5)-3):(nrow(Data5)-99),]   # exclude the last 3 readings (about 30 secs)
HEPAEnd = cpt.meanvar(Data7$PM,method="BinSeg", Q=3, minseglen=8)
Data4$HEPAendnumber = HEPAEnd@cpts[1]        # identify the place where changepoint is
if(Data4$HEPAendnumber==97){                # if there is no changepoint in the last 100 readings then no start HEPA
  Data4$HEPAendtime1 = NA
  Data4$HEPAendtime2 = NA
  Data4$HEPAendvalue1 = NA
  Data4$HEPAendvalue2 = NA
} else {
  Data4$HEPAendtime1 = Data5$Datetime[nrow(Data5)-2-Data4$HEPAendnumber]              # the starttime of end HEPA
  Data4$HEPAendtime2 = Data5$Datetime[nrow(Data5)-3]   # the endtime of end HEPA
  # the mean nephelometer reading in end HEPA
  Data4$HEPAendvalue1 = mean(Data7$PM[1:Data4$HEPAendnumber]) 
  # the mean nephelometer reading after end HEPA period (within the first 100 readings)
  Data4$HEPAendvalue2 = mean(Data7$PM[(Data4$HEPAendnumber+1):97])
}
  
Data4$HEPAstart [Data4$HEPAstvalue1 <= 20 & Data4$HEPAstvalue1 >= -20] <- "Yes"
Data4$HEPAstart [Data4$HEPAstvalue1 >= 20 & Data4$HEPAstvalue1 <= -20] <- "No"

Data4$HEPAEnd [Data4$HEPAendvalue1 <= 20 & Data4$HEPAendvalue1 >= -20] <- "Yes"
Data4$HEPAEnd [Data4$HEPAendvalue1 >= 20 & Data4$HEPAendvalue1 <= -20] <- "No"



```

Start Time | End Time | Run Time | Mean PM | Median PM | Unit ID | Filter ID | SampleID
-----------|----------|-----------|--------|----------|-----------|----------|------
`r as.character(min(indtest_minavg$Time))`  | `r as.character(max(indtest_minavg$Time))` | `r paste(round(nrow(indtest_minavg)/60, digits = 1), "hours")` |`r paste(round(mean(indtest_minavg$PM, na.rm = T), digits = 1), "ug/m^3")` |`r paste(round(median(indtest_minavg$PM, na.rm = T), digits = 1), "ug/m^3")`| `r indtest_minavg$UnitID[1]`| `r indtest_minavg$FilterID[1]`| `r indtest_minavg$SampleID[1]`


Start HEPA|Start HEPA Mean PM | Post Start HEPA Mean PM
:----------:|:--------------:|:----------:|:-------------:
`r Data4$HEPAstart[1]` |`r round(Data4$HEPAstvalue1[1], digits = 3) ` | `r round(Data4$HEPAstvalue2[1], digits = 3) `  



End HEPA | End Mean HEPA PM | Pre End HEPA Mean PM | Low Battery Stop?
:----------:|:-------------:|:-------------:|:-------------:
`r Data4$HEPAEnd[1]`  | `r round(Data4$HEPAendvalue1[1], digits = 3) ` | `r round(Data4$HEPAendvalue2[1], digits = 3) ` | `r as.character(batt$action)`



#Run Time (hours) by Date

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

rows.per.group_date  <- aggregate(rep(1, length(paste0(indtest_minavg$Date))),
                             by=list(indtest_minavg$Date), sum)
names(rows.per.group_date) <- c("Date", "Hours")
rows.per.group_date$Hours <- round(rows.per.group_date$Hours/60, digits = 1)

knitr::kable(rows.per.group_date, caption = "")
```

#Run Time (hours) by 24 hour periods

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}

rows.per.group_date  <- aggregate(rep(1, length(paste0(indtest_minavg.complete$unique_24h))),
                             by=list(indtest_minavg.complete$unique_24h), sum)
names(rows.per.group_date) <- c("24 hour period", "Hours")
rows.per.group_date$Hours <- round(rows.per.group_date$Hours/60, digits = 1)

knitr::kable(rows.per.group_date, caption = "")
```

#Compliance (hours) by Date 

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}


indtest_minavg.complete_d <- ddply(indtest_minavg.complete, .(Date), summarise,  hours_worn = round(sum((compliance_rollmean)/60, na.rm = TRUE), digits = 2))
names(indtest_minavg.complete_d) <- c("Date", "Hours Worn")
knitr::kable(indtest_minavg.complete_d, caption = "")

``` 

#Compliance (hours) by 24 hour periods

```{r, echo=FALSE, message=FALSE, warning=FALSE, results='asis'}


indtest_minavg.complete_d2 <- ddply(indtest_minavg.complete, .(unique_24h), summarise,  hours_worn = round(sum((compliance_rollmean)/60, na.rm = TRUE), digits = 2))
names(indtest_minavg.complete_d2) <- c("Date", "Hours Worn")
knitr::kable(indtest_minavg.complete_d2, caption = "")

``` 

#PM 2.5* timeseries plots

*Plot are the RH-corrected Neph values. There are no HEPA or filter corrections added here.
```{r, echo=FALSE, message=FALSE, warning=FALSE, results='hide', fig.align='center', fig.height= 5, fig.width=7}

ggplot(indtest_minavg, aes(x=Time, y=Flow, color= SampleID)) + geom_point(color = "#00CCFF") + ggtitle("PM2.5")

ggplot(indtest_minavg, aes(x=Time, y=PM, color= SampleID)) + geom_point(color = "#00CCFF") + ggtitle("PM2.5")
ggplot(indtest_minavg, aes(x=Time, y=PM, color= SampleID)) + geom_line(color = "#00CCFF") + ggtitle("PM2.5")

# plot the first 100 readings and show the changepoint
title <-paste("Average during Start HEPA =", round(mean(Data4$HEPAstvalue1), digits = 3 ), "\n", "PM vs Time on 10 second base")
plot(HEPASt,cpt.width=3) + title(main = title,  cex.main = 0.7, col.main = "black") 

# plot the end 100 readings and show the changepoint
title <-paste("Average during End HEPA =", round(mean(Data4$HEPAendvalue1), digits = 3 ), "\n", "PM vs 'Reverse Time' on 10 second base")
plot(HEPAEnd,cpt.width=3) + title(main = title,  cex.main = 0.7, col.main = "black") 


ggplot(indtest_minavg, aes(x=Time, y=Battery, color= SampleID)) + geom_point(color = "#CC33CC") + ggtitle("Battery (V)")
ggplot(indtest_minavg, aes(x=Time, y=RH, color= SampleID)) + geom_point(color = "#666600") + ggtitle("RH (%)")
ggplot(indtest_minavg, aes(x=Time, y=Temp, color= SampleID)) + geom_point() + ggtitle("Temperature (C)")

ggplot(indtest_minavg, aes(x=Time, y=VectorSumComposite, color= SampleID)) + geom_point(color ="#00CC99") + ggtitle("Activity (g)")

ggplot(indtest_minavg, aes(x=Time, y=SD_VectorSumComposite, color= SampleID)) + geom_point(color = "#00CC99") + ggtitle("Standard Deviation of Activity")



```
